Writing a reproducible paper with R Markdown and Pagedown^[Corresponding adress: mail@paulcbauer.eu, camille.landesvatter@mzes.uni-mannheim.de]

### 2021-06-22 paper.Rmd
Added osf link.

---
title: |
  |
  | 
  | \vspace{1cm}Writing a reproducible paper
  | in Pagedown*^[*Corresponding address: mail@paulcbauer.eu. Acknowledgments: I am grateful to all those generous people that invest their time into open-source software.]\vspace{0.5cm}
  |
author: |
  | Paul C. Bauer & Camille Landesvatter
  | Mannheim Centre for European Social Research
date: |
  |       
  | Based on: https://osf.io/q395s
  | This version: `r gsub("^0", "", format(Sys.time(), "%B %d, %Y"))`
  |
  | Feedback? Errors?
  | mail@paulcbauer.eu
  | [Github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)
linestretch: 1.2
colorlinks: true
abstract: \noindent\setstretch{1}The present paper provides a template for a reproducible scientific paper written in R Markdown. Below I outline some of the "tricks"/code (e.g., referencing tables, sections etc.) I had to figure out to produce this document. The underlying files which produce this document can be downloaded [here](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing). I think I got pretty far but there is always room for improvement and more automatization, in parallel to the incredible developments in R and Rstudio (bookdown etc.). I intend to update this file when I discover more convenient code (you can follow any updates through the corresponding [github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)).\vspace{.8cm}
bibliography: references.bib
csl: american-sociological-association.csl
output:
  bookdown::pdf_document2:
    toc: no
    keep_tex: true
mainfont: Times New Roman
sansfont: Times New Roman
fontsize: 12pt
link-citations: true
documentclass: article
geometry: margin=1in
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
# Use chache = TRUE if you want to speed up compilation

# A function to allow for showing some of the inline code
rinline <- function(code){
  html <- '<code  class="r">``` `r CODE` ```</code>'
  sub("CODE", code, html)
}
```

\clearpage

\renewcommand{\baselinestretch}{0.5}\normalsize
\tableofcontents
\renewcommand{\baselinestretch}{1.1}\normalsize

\clearpage

# Why reproducible research (in R)?
Some arguments...

* **Access**: Research is normally funded by taxpayers (researchers are also taxpayers). Hence, it should be freely accessible to everyone without any barriers, e.g., without requiring commercial software. Importantly, researchers from developing countries are even more dependent on free access to knowledge [@Kirsop2005-ro].
* **Reproducability**: Even if you have written a study and analyzed the data yourself you will forget what you did after a few months. A fully reproducible setup will help you to trace back your own steps. Obviously, the same is true for other researchers who may want to understand your work and built on it. It may sound like a joke but why not aim for a document that can be used to reproduce your findings in 500 years. 
* **Errors**: Manual steps in data analysis (e.g., manually copy/pasting values into a table etc.) may introduce errors. R Markdown allows you to **automatize** such steps and/or avoid them.
* **Revisions**: Revising a paper takes much less time if you have all the code you need in one place, i.e., one `.rmd` file. For instance, if you decide to exclude a subset of your data you simply need to insert one line of your code at the beginning and everything is rebuilt/re-estimated automatically.

# Prerequesites
I assume that you are using R on a day-to-day basis. You may have even started to work a little in R Markdown but you don't write your complete paper in R Markdown. If you don't know what R Markdown is watch [this short video](https://vimeo.com/178485416). Then...

* ...install [R](https://www.r-project.org/) and [Rstudio](https://www.rstudio.com/) (most recent versions) [@R2017; @Rstudio2015].
* ...install [tinytex](https://yihui.name/tinytex/), a lightweight version of Tex Live [@markdown2017; @tinytex].
```{r eval=FALSE, include=TRUE}
install.packages(c('tinytex', 'rmarkdown'))
tinytex::install_tinytex()
```
* ...install the packages below using the code below [@bookdown1; @bookdown2; @knitr1; @knitr2; @knitr3; @kableextra; @plotly].
```{r eval=FALSE, include=TRUE}
install.packages(c("rmarkdown", "knitr", "kableExtra",
                   "stargazer", "plotly", "knitr",
                   "bookdown"))
```
* ...download the 5 input files I created --- `paper.rmd`, `references.bib`, `data.csv` and `american-sociological-association.csl` --- from [this folder](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing). Ignore the other files.

* ...learn R and read about the other underlying components namely [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Latex](https://en.wikipedia.org/wiki/LaTeX).

# Basics: Input and output files
All the files you need to produce the present PDF file are the input files...

* ...a `paper.rmd` file (the underlying R Markdown file).
* ...a `references.bib` file (the bibliography).
    + I use paperpile to manage my references and export the `.bib` file into the folder that contains my `.rmd` file.
* ...a `data.csv` file (some raw data).
* ... a `american-sociological-association.csl` file that defines the style of your bibliography.^[You can download various citation style files from this webpage: https://github.com/citation-style-language/styles.]

[Download these files](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing) and save them into a folder. Close R/Rstudio and directly open `paper.rmd` with RStudio. Doing so assures that the working directory is set to the folder that contains `paper.rmd` and the other files.^[You can always check your working directory in R with `getwd()`.]

Once you run/compile the `paper.rmd` file in Rstudio it creates mainly two output files:

* `paper.tex`
* `paper.pdf` (the one you are reading right now)

In addition, there may be files that you generate and store locally in the folder during the compilation process. This is the case for some of the Plotly graphs below.

Ideally, we can simply provide others with a `zip` folder that contains both our input files and our output files. Then it's possible to reproduce the process from managing/analyzing some raw data to producing the final scientific article.

Below we always display the R code in the chunks that produce the output. In your paper you will normally only present outputs (e.g., tables, figures etc.) by choosing the chunk option "Show output only" in R Studio. The chunk commands itself are not displayed but they do matter for referencing etc. So simply orient yourself at the underlying `paper.rmd` file.

# Referencing within your document

@@

# Software versioning
Software changes and gets updated, especially with an active developer community like that of R. Luckily you can always access [old versions of R](https://cran.r-project.org/bin/windows/base/old/) and old version of R packages in [the archive](https://cran.r-project.org/src/contrib/Archive/). In the archive you need to choose a particular package, e.g dplyr and search for the right version, e.g., `dplyr_0.2.tar.gz`. Then insert the path in the following function: `install.packages("https://....../dplyr_0.2.tar.gz", repos=NULL, type="source")`. Ideally, however, results will be simply reproducible in the most current R and package versions.

@@

# Data
## Import
```{r, echo=TRUE}
data <- read.csv("data.csv")
head(data)
```

@@

# Tables {#sec:tables}
Producing good tables and referencing these tables within a R Markdown PDF has been a hassle but got much better. Examples that you may use are shown below. The way you reference tables is slightly different, e.g., for `stargazer` the label is contained in the function, for `kable` it's contained in the chunk name.

@@

Table \@ref(tab1) shows summary stats of your data.^[To reference the table where you set the identifier in the stargazer function you only need to use the actual label, i.e., ´tab1´.] I normally use `stargazer()` [@hlavac2013stargazer] which offers extreme flexibility regarding table output (see `?stargazer`).

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
library(stargazer)
stargazer(cars, 
          title = "Summary table with stargazer",
          label="tab1", 
          table.placement = "H", 
          header=FALSE)
```

@@

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
library(stargazer)
model1 <- lm(speed ~ dist, data = cars)
model2 <- lm(speed ~ dist, data = cars)
model3 <- lm(dist ~ speed, data = cars)
stargazer(model1, model2, model3,
          title = "Regression table with stargazer",
          label="tab2", 
          table.placement = "H", 
          column.labels = c("M1", "M2", "M3"),
          model.numbers = FALSE,
          header=FALSE)
```

@@

Another great function is `kable()` (`knitr` package) in combination with `kableExtra`. Table \@ref(tab:tab3) provides an example.^[To reference the table produced by the chunk you need to add ´tab:´ to the chunk name, i.e., ´tab:tab3´.] Again you can modify so many things in both the `kable()` and the `kable_styling()` function. See [this overview](https://haozhu233.github.io/kableExtra/awesome_table_in_pdf.pdf) of all the kable stylings that are possible provided by the package author himself.

```{r tab3, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
library(knitr)
library(kableExtra)
kable(cars[1:10,], row.names = TRUE, 
      caption = 'Table with kable() and kablestyling()', 
      format = "latex", booktabs = T) %>%
        kable_styling(full_width = T, 
                      latex_options = c("striped", 
                                        "scale_down",
                                        "HOLD_position"),
                      font_size = 10)
```

# Inline code & results
Reproduction reaches new heights when you work with inline code. For instance, you can automatize the display of certain coefficients within the text. An example is to include estimates, e.g., the coefficient of `dist` of the model we ran above. `r rinline("round(coef(model1)[2], 2)")` will insert the coefficient as follows: `r round(coef(model1)[2], 2)`. Or `r rinline("3 + 7")` will insert a `r 3 + 7` in the text.      
Inline code/results that depend on earlier objects in your document will automatically be updated once you change those objects. For instance, imagine a reviewer asks you to omit certain observations from your sample. You can simply do so in the beginning of your code and push play subsequently.. at time you might have to set `cache = FALSE` at the beginning so that all the code chunks are rerun.    
Researchers often avoid referring to results in-text etc. because you easily forget to change them when revising a manuscript. At the same it can make an article much more informative and easier to read, e.g., if you discuss a coefficient in the text you can directly show it in the section in which you discuss it. Inline code allows you to do just that. R Markdown allows you to that do so in a reproducible and automatized manner.

@@

@@

* Keep all files of your project (that matter for producing the PDF) in one folder without subfolders. You can zip and directly upload that folder to the [Harvard dataverse](https://dataverse.harvard.edu/).
* Make sure that filenames have a logic to them.
    + Main file with text/code: "paper.rmd", "report.rmd" 
    + Data files: "data_xxxxxx.*"
    + Image files: "fig_xxxxxx.*"
    + Tables files: "table_xxxx.*"
    + etc.
    + Ideally, your filenames will correspond to the names in the paper. For instance, Figure 1 in the paper may have a corresponding file called `fig_1_xxxxx.pdf`.
* Use the document outline in R studio (Ctrl + Shift + O) when you work with R Markdown.
* Name rchunks according to what they do or produce:
    + "`fig-...`" for chunks producing figures
    + "`table-...`" for chunks producing tables
    + "`model-...`" for chunks producing model estimates
    + "`import-...`" for chunks importing data
    + "`recoding-...`" for chunks in which data is recoded
* Ideally give chunks the labels of the figures/tables in the final publication, e.g., "`fig-1`" is the name of the chunk that produces Figure 1 in the final paper.
* Use "really" informative variable names:
    + Q: What do you think does the variable *trstep* measure? It actually measures trust in the European parliament. 
        + How could we call this variable instead? Yes, `trust.european.parliament` which is longer but will probably be understood by another researcher. 
    + If your setup is truly reproducible you will probably re-use the variable names that you generate as variable names in the tables you produce. Hence, there is an incentive to use good names.
* Use unique identifiers in the final document:
    + e.g., name the models you estimate "M1", "M2" etc.
    + These unique names should also appear in the published paper.
    + Think of someone who wants to produce Figure 1/Model 1 in your paper but doesn't find it in your code...

@@

# References

\linespread{1}

@@

\clearpage

\appendix
\addtocontents{toc}{\protect\setcounter{tocdepth}{2}}

\renewcommand{\thesection}{A}

\setcounter{page}{1}

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\renewcommand{\figurename}{Table}

\setcounter{figure}{0}
\renewcommand\thefigure{A\arabic{figure}}
\renewcommand{\figurename}{Figure}

\clearpage
\pagenumbering{gobble}

\vspace*{7cm}

\begin{center}
\begin{huge}
Online Appendix
\end{huge}
\end{center}
\vspace{3cm}

\clearpage
\pagenumbering{arabic}

# Online appendix

>>>>

---
title: "Writing a reproducible paper with R Markdown and Pagedown^[Corresponding adress: mail@paulcbauer.eu, camille.landesvatter@mzes.uni-mannheim.de]"
shorttitle: "Writing a reproducible paper with R Markdown and Pagedown"
author:
  - name: Paul C. Bauer
    affiliation: Mannheim Centre for European Social Research
    email: <mail@paulcbauer.eu>
  - name: Camille Landesvatter
    affiliation: Mannheim Centre for European Social Research
    email: <camille.landesvatter@mzes.uni-mannheim.de>
date: 'First version: 20 June, 2021<br>This version: `r format(Sys.time(), "%d %B, %Y")`'
download-link: 'Download: https://osf.io/k8jhx'
abstract: 'The present paper provides a template to write a reproducible scientific paper with R Markdown and Pagedown.^[Based on an earlier R Markdown template that can be downloaded under https://osf.io/q395s [see @Bauer2018-dl].] Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document. The underlying files which produce this document can be downloaded^[https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing]. Importantly, we also provide different CSS and HTML files that can be used to achieve a pdf output with the look of a "working paper". We are convinced that in the future there will be many improvements and developments with regards to RStudio, R markdown and R pagedown. We intend to update this file when we discover more convenient code (you can follow any updates through the corresponding [github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_pagedown/)).'
# at least one keyword must be supplied
keywords: [open science, transparency, replication, R, markdown, pagedown]
links-to-footnotes: true
paged-footnotes: true
bibliography: references.bib
output: 
  pagedown::jss_paged:
    template: wp_paged.html
    self_contained: true
    css: ['wp.css', 'wp-fonts.css', 'wp-page.css']
    csl: american-sociological-association.csl
# uncomment the next line if you only want to knit to HTML and not PDF
knit: pagedown::chrome_print
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
# Use chache = TRUE if you want to speed up compilation
# A function to allow for showing some of the inline code
rinline <- function(code){
  html <- '<code  class="r">``` `r CODE` ```</code>'
  sub("CODE", code, html)
}
```

\newpage
# Why reproducible research (in R)?

Some arguments...

* **Access**: Research is normally funded by taxpayers (researchers are also taxpayers). Hence, it should be freely accessible to everyone without any barriers, e.g., without requiring commercial software. Importantly, researchers from developing countries are even more dependent on free access to knowledge [@Kirsop2005-ro].

* **Reproducability**: Even if you have written a study and analyzed the data yourself you will forget what you did after a few months. A fully reproducible setup will help you to trace back your own steps. Obviously, the same is true for other researchers who may want to understand your work and built on it. It may sound like a joke but why not aim for a document that can be used to reproduce your findings in 500 years.

* **Errors**: Manual steps in data analysis (e.g., manually copy/pasting values into a table etc.) may introduce errors. R Markdown allows you to **automatize** such steps and/or avoid them.

* **Revisions**: Revising a paper takes much less time if you have all the code you need in one place, i.e., one `.rmd` file. For instance, if you decide to exclude a subset of your data you simply need to insert one line of your code at the beginning and everything is rebuilt/re-estimated automatically.

# Why Pagedown?

Formatting text as PDF is probably one of the most widespread standards in the scientific community, especially when it comes to submitting papers and similar documents. The traditional way to well-formatted and good-looking PDFs is often through LaTeX or Word.

The fairly new `pagedown` R package takes a completely new approach. While the main purpose of pagedown is to create high-quality PDFs, the idea is to take advantage of modern web technologies (HTML/JSS/Javascript) with which one can design web pages and eventually print those to PDF.

While web pages are usually single-page scrollable documents, pagedown uses the JavaScript library `Paged.js` which allows documents to be paginated with elements like headers, footers and everything a readable scientific paper will need. Additionally, pagedown documents are based on R Markdown.

# Prerequesites

We assume that you are using R on a day-to-day basis and you may have even started to work in R Markdown. If you don't know what R Markdown is there are many great resources (e.g. watch this [short video](https://vimeo.com/178485416)). Also, there is a template for a scientific paper written in R markdown by one of us two (Paul), which follows the same idea and structure just as this review ([github repo](https://github.com/paulcbauer/Writing_a_reproducable_paper_in_rmarkdown/)). The template compiles information, operations, tips and tricks you will very likely need to create a well-formatted scientific paper with R markdown.

Based on R Markdown, Pagedown allows you to create custom and well-formatted (paged) HTML Documents. For a comprehensive overview watch [this video](https://www.rstudio.com/resources/rstudioconf-2019/pagedown-creating-beautiful-pdfs-with-r-markdown-and-css/) which is a record of a talk introducing `pagedown` given by Yihui Xie (who in addition to Romain Lesur developed the `pagedown` package). If you are not in a video watching mood find the slides [here](https://slides.yihui.org/2019-rstudio-conf-pagedown.html#1).

Then...

* ...install [R](https://www.r-project.org/) and [Rstudio](https://www.rstudio.com/) (most recent versions) [@R2017; @Rstudio2015].
* ...install the "pagedown"-package from github using the code below.
```{r eval=FALSE, include=TRUE}
remotes::install_github('rstudio/pagedown')
```
* ...also install the packages below using the code below [@bookdown1; @bookdown2; @knitr1; @knitr2; @knitr3; @kableextra; @plotly].
```{r eval=FALSE, include=TRUE}
install.packages(c("rmarkdown", "knitr", "kableExtra",
                   "stargazer", "modelsummary", "plotly", "knitr"))
```
* ...download the 4 input files we created --- `paper.rmd`, `references.bib`, `data.csv` and `american-sociological-association.csl` --- from [this folder](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing). Ignore the other files.
* ...also download the 4 styling files we created: `wp_paged.html`, `wp.css`, `wp-fonts.css` and `wp-pages.css`.
* ...store all 8 files from above together in one folder (and use this folder as your working directory later on)
* ...learn R and read about the other underlying components namely [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Latex](https://en.wikipedia.org/wiki/LaTeX).
* ...pagedown comes with several Rmd-templates (presentations, poster, thesis, etc.) and via this review we provide another template for a working paper style. If however you want to modify single aspects or create your own template, you will need to at least gain some basic skills in [CSS](https://www.w3schools.com/css/) and [HTML](https://www.w3schools.com/html/).

# Basics: Input files, output files and the YAML header

All the files you need to produce the present PDF file are:

1. the input files:

* `paper.rmd` (the underlying R Markdown file).
* `references.bib` (the bibliography).
    + I use paperpile to manage my references and export the `.bib` file into the folder that contains my `.rmd` file.
* `data.csv` (some raw data).
* `american-sociological-association.csl` (defines the style of your bibliography).^[You can download various citation style files from this webpage: https://github.com/citation-style-language/styles.]

2. the "styling" files:

Basically, these are files you will need to specify in the YAML of your rmd-file, so that R and ultimately pagedown recognizes the certain style you want to achieve for your document. With using our templates, you will create a document that has the "look" of a working paper.

* `wp_paged.html`
* `wp.css`
* `wp-fonts.css`
* `wp-pages.css`

Take `paper.rmd` (the underlying R Markdown file of this pdf) and have a look at the YAML (line \#18 - \#22) to see how to specifiy these files.
Basically, what happens here is that within the [jss_paged function](https://rdrr.io/cran/pagedown/man/jss_paged.html) we additionally specify that we want to use custom CSS and custom HTML.

[Download these files](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing) and save them into a folder. Close R/Rstudio and directly open `paper_pagedown.rmd` with RStudio. Doing so assures that the working directory is set to the folder that contains `paper.rmd` and the other files.^[You can always check your working directory in R with `getwd()`.]

Once you run/compile the `paper.rmd` file in Rstudio it creates a output file called `paper_pagedown.html`.

By using pagedown's `chrome_print` function in the YAML (line \#25) your html based web page will be printed to `paper_pagedown.pdf` (the one you are reading right now).

Both outputs will be saved in your working directory.

\newpage
# Referencing within your document

@@

# Software versioning

Software changes and gets updated, especially with an active developer community like that of R. Luckily you can always access [old versions of R](https://cran.r-project.org/bin/windows/base/old/) and old version of R packages in [the archive](https://cran.r-project.org/src/contrib/Archive/). In the archive you need to choose a particular package, e.g dplyr and search for the right version, e.g., `dplyr_0.2.tar.gz`. Then insert the path in the following function: `install.packages("https://....../dplyr_0.2.tar.gz", repos=NULL, type="source")`. Ideally, however, results will be simply reproducible in the most current R and package versions.

@@

# Data
## Import

Generally, code is evaluated by inserting regular `R Markdown` blocks.

```{r}
x <- 1:10
x
```

Below we import an exemplary dataset ([download](https://drive.google.com/drive/folders/1zJP3cNPrHN-gj0rcmbHQgg-XA0hqDXdd?usp=sharing)).

```{r, echo=T, results="raw"}
data <- read.csv("data.csv")
head(data)
```

@@

# Tables {#sec:tables}

Producing good tables and referencing these tables within a R Markdown PDF has been a hassle but got much better. Examples that you may use are shown below. The way you reference tables is slightly different, e.g., for `stargazer` the label is contained in the function, for `kable` it's contained in the chunk name.

@@

I normally use `stargazer()` [@hlavac2013stargazer] which offers extreme flexibility regarding table output (see `?stargazer`).

### Summary table with stargazer

Table \@ref(tab:summary) shows summary stats of your data.

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
library(stargazer)
stargazer(cars, 
          title = "Summary table with stargazer",
          label="tab:summary", 
          table.placement = "H", 
          type="html",
          header=FALSE)
```

### Regression table with stargazer

@@

```{r, eval=TRUE, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
library(stargazer)
model1 <- lm(speed ~ dist, data = cars)
model2 <- lm(speed ~ dist, data = cars)
model3 <- lm(dist ~ speed, data = cars)

stargazer(model1, model2, model3,
          label="tab2",
          table.placement = "H", 
          column.labels = c("M1", "M2", "M3"),
          type="html",
          model.numbers = FALSE,
          header=FALSE)
```

@@

Another great function is `kable()` (`knitr` package) in combination with `kableExtra`. Table \@ref(tab:tab3) provides an example.^[To reference the table produced by the chunk you need to add ´tab:´ to the chunk name, i.e., ´tab:tab3´.]

```{r tab3, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
library(knitr)
library(kableExtra)

kable(mtcars[1:10,], row.names = TRUE, 
      caption = 'Table with kable() and kablestyling()', 
      format = "html", booktabs = T) %>%
        kable_styling(full_width = T, 
                      latex_options = c("striped", 
                                        "scale_down",
                                        "HOLD_position"),
                      font_size = 10)

```

## modelsummary

The `modelsummary` package provides a variety of tables and plots to summarize statistical models and data in R. Modellsummary plots and tables are highly customizable and they can be saved to almost all formats, e.g., HTML, PDF and Markdown. This makes ist especially easzz to embed them in dznmaic documents.

Please look at the package's extensive [documentation](https://vincentarelbundock.github.io/modelsummary/index.html) where they also show examples for almost any plot or table you might be looking for.

In this template we demonstrate an example for modelsummary's `datasummary` function. `Datasummary` creates frequency tables, crosstab tables, correlation tables, balance tables and many **more**.

### Summarize numeric variables with modelsummary

Table \@ref(tab:tab4) shows a summary table for numeric variables.

```{r tab4, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
library(modelsummary)

mtcars_df <- mtcars
mtcars_df$vs <- as.logical(mtcars_df$vs)
mtcars_df$cyl <- as.factor(mtcars_df$cyl)
datasummary_skim(mtcars, type="numeric", output="html", histogram=T)
```

### Summarize categorical variables with modelsummary

Table \@ref(tab:tab5) shows a summary table for categorical variables.

```{r tab5, echo=TRUE, message=FALSE, warning=FALSE, results="asis"}
datasummary_skim(mtcars_df, type="categorical", output="html")
```

# Inline code & results

Reproduction reaches new heights when you work with inline code. For instance, you can automatize the display of certain coefficients within the text. An example is to include estimates, e.g., the coefficient of `dist` of the model we ran above. `r rinline("round(coef(model1)[2], 2)")` will insert the coefficient as follows: `r round(coef(model1)[2], 2)`. Or `r rinline("3 + 7")` will insert a `r 3 + 7` in the text.      
Inline code/results that depend on earlier objects in your document will automatically be updated once you change those objects. For instance, imagine a reviewer asks you to omit certain observations from your sample. You can simply do so in the beginning of your code and push play subsequently.. at time you might have to set `cache = FALSE` at the beginning so that all the code chunks are rerun.    
Researchers often avoid referring to results in-text etc. because you easily forget to change them when revising a manuscript. At the same it can make an article much more informative and easier to read, e.g., if you discuss a coefficient in the text you can directly show it in the section in which you discuss it. Inline code allows you to do just that. R Markdown allows you to that do so in a reproducible and automatized manner.

@@

## Interactive graphs

# Compiling

To view your paper, pagedown requires a web server (since it is based von paged.js)^[open-source library to paginate content in the browser]. By compiling a document, R Studio will display your HTML page through a local web server, i.e., paged.js will work in RStudio Viewer.

There are several options, depending on your intention:

- click on the `Knit` button which by default will provide a HTML document in the RStudio viewer pane (the HTML will be stored in your working directory)
- use pagedown's `chrome_print` function in the YAML (uncomment line \#24 of this `Rmd` file) if additionally you want your HTML based web page to be printed to be PDF (the PDF will be stored in your working directory)
- to "live"-preview your pages do not click on the `Knit` button but use the **xaringan** [@xaringan] RStudio add-in _Infinite Moon Reader_. You can simply call the function `xaringan::inf_mr()` (within your console). This will launch a local web server via the **servr** package (Xie 2021a) and display your pages in the RStudio viewer. Each time you save your document (_Ctrl+S_) xaringan updates your pages in the viewer.

- If you use the option `self_contained: false` (see line \#21 of this `Rmd` file) (change to true for a self-contained document, but it'll be a litte slower for Pandoc to render), don't click on the `Knit` button in RStudio. Use instead the **xaringan** [@xaringan] RStudio add-in _Infinite Moon Reader_.

@@

* Keep all files of your project (that matter for producing the PDF) in one folder without subfolders. You can zip and directly upload that folder to the [Harvard dataverse](https://dataverse.harvard.edu/).
* Make sure that filenames have a logic to them.
    + Main file with text/code: "paper.rmd", "report.rmd" 
    + Data files: "data_xxxxxx.*"
    + Image files: "fig_xxxxxx.*"
    + Tables files: "table_xxxx.*"
    + etc.
    + Ideally, your filenames will correspond to the names in the paper. For instance, Figure 1 in the paper may have a corresponding file called `fig_1_xxxxx.pdf`.
* Use the document outline in R studio (Ctrl + Shift + O) when you work with R Markdown.
* Name rchunks according to what they do or produce:
    + "`fig-...`" for chunks producing figures
    + "`table-...`" for chunks producing tables
    + "`model-...`" for chunks producing model estimates
    + "`import-...`" for chunks importing data
    + "`recoding-...`" for chunks in which data is recoded
* Use "really" informative variable names:
    + Q: What do you think does the variable *trstep* measure? It actually measures trust in the European parliament. 
        + How could we call this variable instead? Yes, `trust.european.parliament` which is longer but will probably be understood by another researcher. 
    + If your setup is truly reproducible you will probably re-use the variable names that you generate as variable names in the tables you produce. Hence, there is an incentive to use good names.
* Use unique identifiers in the final document:
    + e.g., name the models you estimate "M1", "M2" etc.
    + These unique names should also appear in the published paper.
    + Think of someone who wants to produce Figure 1/Model 1 in your paper but doesn't find it in your code...

@@

\newpage

# References {-}

@@

<!--  to generate a citation entry for BibTeX, you can pass the returned object of citation() to toBibtex(), e.g. toBibtex(citation("xaringan")). Then  copy the output to a .bib file and add a unique citation key -->
<!-- https://bookdown.org/yihui/rmarkdown-cookbook/write-bib.html -->

# Online appendix {-}
